# Reproducible Research - Peer Assessment 1

Set global chunk options

```{r set chunk options, echo = TRUE}
opts_chunk$set(echo = TRUE, results = "show", cache = TRUE)
```

## Loading and preprocessing the data

```{r load data}
csvfile <- "activity.csv"
adf <- read.csv(csvfile, header = TRUE)
adfc <- adf[complete.cases(adf),]
````

## What is mean total number of steps taken per day?

- first calculate total steps by day *stepsbyday*
- histogram the total steps taken each day

```{r histogram}
stepsbyday <- aggregate(steps ~ date, data = adfc, FUN = sum)
hist(stepsbyday$steps, 
     breaks = 7,
     col = "green", 
     main = "Histogram of Steps Taken Each Day", 
     xlab ="steps by day") 
```

mean and mediam

```{r}
adfmean <- mean(adfc$steps)
adfmedian <- median(adfc$steps)
```

The **steps' mean** is **`r adfmean`**  
The **steps' median** is **`r adfmedian`**  

## What is the average daily activity pattern?

```{r}
stepsbyint <- aggregate(steps ~ interval, data = adf, FUN = mean)
plot(stepsbyint$interval, stepsbyint$steps, main="Mean of Steps by Interval", xlab="5 minute interval", ylab = "Mean of Steps", type="l")
maxinterval <- stepsbyint[stepsbyint$steps==max(stepsbyint$steps),]
```

**`r maxinterval$interval`** is the 5 minute interval which contains the maximum number of steps on the average 

## Imputing missing values

The number of missing values is
```{r}
sum(is.na(adf))
```

Each NA value in steps is filled up with the appropriate mean of steps depending on the interval using *stepsbyint* data frame previously calculated
```{r}
adff <- adf
for (i in 1:nrow(adff)) {
    if (is.na(adff[i, "steps"])) {
       adff[i, "steps"] <- stepsbyint[stepsbyint$interval == adff[i, "interval"], "steps"]       
    }
}
```

check there are not missing values in the new data frame

```{r}
sum(is.na(adff))
```

 
```{r histogram filled up na}
stepsbydayf <- aggregate(steps ~ date, data = adff, FUN = sum)
hist(stepsbydayf$steps, 
     breaks = 7,
     col = "green", 
     main = "Histogram of Steps Taken Each Day, filled up dataset", 
     xlab ="steps by day") 
```

mean and mediam for the filled up dataset, and the differences with the first part of assignment

```{r}
adffmean <- mean(adff$steps)
adffmedian <- median(adff$steps)
meandif <- adfmean - adffmean 
mediandif <- adfmedian - adffmedian
```
The **steps' mean** is **`r adffmean`**, the difference is **`r meandif`**  
The **steps' median** is **`r adffmedian`**, the difference is **`r mediandif`**

## Are there differences in activity patterns between weekdays and weekends?

```{r weekday and weekend}
Sys.setlocale("LC_TIME", "English")

adff$weekday <- weekdays(as.Date(adff$date))

for (i in 1:nrow(adff)) {
    if ( (adff[i, "weekday"]) %in% c("Saturday","Sunday") ) {
       adff[i, "weekday"] <- "weekend"
       }
    else {
       adff[i, "weekday"] <- "weekday"       
       }
            
}
```

```{r steps sum aggrate by inteval and weekday }
adffag <- aggregate(list(steps=adff$steps),list(interval=adff$interval,weekday=adff$weekday),mean)

library(lattice)
xyplot(steps ~ interval | weekday, data = adffag, 
       layout = c(1, 2), type = "l", ylab = "Number of steps", main = "Average Steps")


